\textsf{Are frontier models superhuman chemists?}

### 2024-03-12 src/tex/appendix.tex
add some more analysis leads

\begin{figure}
    \centering
    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\cite{flesch1948new} is a measure of how easy a text is to read. It is calculated based on the average number of syllables per word and the average number of words per sentence. The higher the score, the easier the text is to read. The distribution of the scores of the questions is shown in the histogram. The mean score is \variable{output/flesch_kincaid_mean.txt} and the median score is \variable{output/flesch_kincaid_median.txt}.}
    \label{fig:flesch_kincaid_reading_ease}
\end{figure}

@@

\subsection{Model performance}

@@

\begin{figure}
    \centering
    \includegraphics{figures/experience_vs_correctness.pdf}
    \script{analyze_human_data.py}
    \label{fig:experience_vs_correctness}
    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers (Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\cite{flesch1948new} is a measure of how easy a text is to read. It is calculated based on the average number of syllables per word and the average number of words per sentence. The higher the score, the easier the text is to read. The distribution of the scores of the questions is shown in the histogram. }
    \label{fig:flesch_kincaid_reading_ease}
\end{figure}

@@

\subsection{Model performance}
We also evaluated the model performance on the entire chembench corpus. 
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models.

\begin{figure}
    \label{fig:barplot_all_correct_all_questions}
\end{figure}

\Cref{fig:all_questions_models_completely_correct_radar} shows the performance of the models on the different topics of the chembench corpus.

@@

\begin{figure}
    \centering
    \includegraphics{figures/experience_vs_correctness.pdf}
    \script{analyze_human_data.py}
    \label{fig:experience_vs_correctness}
    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
\end{figure}

### 2024-03-12 src/tex/ms.tex
update text
add some more analysis leads

Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
Most of these insights are in form of text in scientific publications. 
Thus, operating on such text might be our best way of unlocking and \emph{learning} from these insights.
This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020}
However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards}
To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).

@@

Because many widely utilized systems do only provide access to text completions (and not the raw model outputs), chembench is designed to operate on text completions.
This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are conventionally used for model evaluations) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).

To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\cite{Huggingface} on the chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. In this figure, we show the percentage of questions that the models answered completely correctly.
We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance with vertical lines.
The figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric.
Many other models also outperform the average human performance.
% Bar chart

@@

Given the large interest in tool-augmented systems, the bad performance of these systems in our benchmark is striking. 
Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
With the default ReAct setup, however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.

% Radar chart
To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified the questions into these topics.
We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.

\begin{figure}
    \centering
    \includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
    \caption{Caption}
    \label{fig:enter-label}
\end{figure}

@@

% Estimate of difficulty
One might wonder whether the models are capable of estimating if they will be able to answer a question correctly.
To investigate this, we prompted some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.

@@

>>>>

Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
Most of these insights are in form of text in scientific publications. 
Thus, operating on such text might be our best way of unlocking and \emph{learning} from these insights.
This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).

@@

Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), chembench is designed to operate on text completions.
This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\cite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).

To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\cite{Huggingface} on the chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
In this figure, we show the percentage of questions that the models answered completely correctly.
We use horizontal bars to indicate the performance of various models and highlight statistics (e.g., the average) of the human performance with vertical lines.
Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly exceeds the average performance of our human scores.
Many other models also outperform the average human performance. It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
% Bar chart

@@

Given the large interest in tool-augmented systems, the bad performance of these systems in our benchmark is striking. 
Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
With the default tool augmented setup (using the so-called ReAct method), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.

% Radar chart
To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified the questions into these topics based and hand-crafted rules as well as classifier models.
We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.

\begin{figure}
    \centering
    \includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
    \caption{Caption}
    \label{fig:performance_by_topic}
\end{figure}

In this spider chart, the worst score for every dimension is 0 (no question answered completely correctly) and the best score is 1 (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models. 
While this question is relatively easy () for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound. 
This is mirrored in the questions about point groups of compounds ...

For all topics in chembench, proprietary models outperform their open counterparts.

% Can we say something how this links to changes we have to make to the training data

@@

\begin{figure}
    \centering
    %\includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
    \caption{Caption}
    \label{fig:calculation_performance}
\end{figure}

% Estimate of difficulty
One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
To investigate this, we prompted some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.

@@

\subsection{Confidence estimate}
To estimate the models' confidence we prompted them with the question (and answer options for \gls{mcq}) and the task to rate their confidence to produce the right answer on a scale from 1 to 5.
